Quantum Walks and Applications to Quantum Money

### 2025-04-01 presentation.tex
new commit
presentation updated

% Packages
\usepackage{graphicx}    % For including images
\usepackage{amsmath}     % For mathematical symbols
\usepackage{braket}      % For Dirac notation
\usepackage[absolute,overlay]{textpos}
\usepackage{amsmath}
% \usepackage{amsmath, amssymb, dsfont}
% \usepackage{algorithm}
% \usepackage{algorithmicx}
% \usepackage{enumitem}

@@

% \begin{frame}{Program and Courses Taken}

%     \begin{textblock*}{12cm}(0.5cm,1.5cm)
%        \begin{itemize}
%         \item Program: M.A.Sc in Software Engineering
%         \vspace{1cm}
%         \item CAS 701,  Logic \& Discrete Mathematics
%         \item COMPSCI 6TE3, Continuous optimization
%         \item CAS 721, Combinatorics \& Computing
%         \item CAS 741, Development of Scientific Computation Software
%        \end{itemize}
%     \end{textblock*}

%     \title{SDSD}

% \begin{frame}{Seminars}

%     \begin{textblock*}{12cm}(0.5cm,1.5cm)
%        \begin{itemize}
%         \item I have participated in 6 seminars during my program.
%        \end{itemize}
%     \end{textblock*}

% \begin{frame}{Poster}

%     \begin{figure}
%         \centering
%         \includegraphics[width=1\textwidth]{poster.png}
%         \caption{Presented at Nov. 6, 2024}
%     \end{figure}

@@

@@

@@

@@

@@

@@

@@

\begin{frame}{The cmpIndex Algorithm}

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        Given a state $\ket{G^{(h)} * x}$, there is an efficient algorithm for computing $h$.
        Specifically, there is a unitary operator that performs the transformation $\ket{G^{(h)} * x} \ket{0} \mapsto \ket{G^{(h)} * x} \ket{h}$:

        \[
        \ket{G^{(h)} * x}\ket{0} \mapsto \ket{G^{(h)} * x} \frac{1}{\sqrt{|G|}}\sum_{k \in G}\ket{k}
        \]
        And then apply the unitary $\sum_{k \in G}U_k \otimes \ket{k}\bra{k}$:

        \[
        \mapsto \frac{1}{\sqrt{|G|}}\sum_{k \in G}\ket{G^{(h)}*x}\chi(-k,h)\ket{k}
        \]

        Finally, applying the inverse quantum Fourier transform to the second register yields:

        \[
           \mapsto  \ket{G^{(h)}*x}\ket{h}
        \]
    \end{textblock*}

@@

        % \begin{align*}
        %     2\ket{0} \bra{0} \otimes TT^* - 1
        %     & = 2\ket{0} \bra{0} \otimes \sum_{y \in X} \sum_{b \in \{0, 1\}} \ket{y, b} \ket{\phi_{yb}} \bra{y, b} \bra{\phi_{yb}} - 1 \\

        %     & = U_T (2 \ket{0} \bra{0} \otimes 1_{X, b} \otimes \ket{0, 0} \bra{0, 0} - 1) U_T^*.

        % \end{align*}

        \[
        2\ket{0} \bra{0} \otimes TT^* - 1 = 2\ket{0} \bra{0} \otimes \sum_{y \in X} \sum_{b \in \{0, 1\}} \ket{y, b} \ket{\phi_{yb}} \bra{y, b} \bra{\phi_{yb}} - 1 
        \]

        \[
        = 2\sum_{y \in X} \sum_{b \in \{0, 1\}} \ket{0} \ket{y, b} \ket{\phi_{yb}} \bra{0} \bra{y, b} \bra{\phi_{yb}} - 1 
        \]

        \[
        = U_T \Big( 2\sum_{y \in X} \sum_{b \in \{0, 1\}} \ket{0} \ket{y, b} \ket{0, 0} \bra{0} \bra{y, b} \bra{0, 0} - 1 \Big) U_T^* 
        \]

        \[
        = U_T (2 \ket{0} \bra{0} \otimes 1_{X, b} \otimes \ket{0, 0} \bra{0, 0} - 1) U_T^*
        \]

        Now, for any state $\ket{\psi}$:

        \[
        (2\ket{0}\bra{0}\otimes TT^* - 1)\ket{0}\ket{\psi} = \ket{0}(2TT^*-1)\ket{\psi}
        \]

\begin{frame}{Group Action Quantum Walks}

        Given a regular group action $(G, X, *)$ with a fixed element $x \in X$ and a set  $Q = \{q_1, q_2, \dots, q_k\} \subset G$, let $\Gamma = (X, E)$ be a graphs with vertex set $X$ and edge set consisting of pairs $(x, y) \in X \times X$ such that $y = q * x$ for some $q \in Q$. The adjacency matrix of $\Gamma$ is
        \[ A = \sum_{h \in G} \lambda_h \ket{G^{(h)} * x} \bra{G^{(h)} * x}, \]
        where:
        \begin{itemize}
            \item $\lambda_h = \sum_{q \in Q} \chi(h, q)$
            \item the eigenvectors $\ket{G^{(h)} * x}$ depend only on $G$
        \end{itemize}

@@

        \begin{itemize}
        \item $\ver(h, \ket{\psi})$. First, check whether $\ket{\psi}$ has support in $X_\lambda$. If not, return $0$. Then, apply $\comph$ to the state $\ket{\psi}\ket{0}$, and measure the second register to obtain some $h' \in G_\lambda$. If $h' = h$, return $1$; otherwise return $0$.
        \end{itemize}

@@

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        Lemma: The money state $\ket{\mathbb{Z}_N^{(h)} * x}_H$ is an eigenstate of $W$ with eigenvalue $e^{i\lambda_h t}$.
        % \vspace{1cm}

        \textit{Proof:}
        \begin{align*}
            e^{iAt} \ket{\mathbb{Z}_N^{(h)} * x}_H
            & = \sum_{g \in \mathbb{Z}_N} e^{i\lambda_gt} \ket{\mathbb{Z}_N^{(g)} * x}\braket{\mathbb{Z}_N^{(g)} * x|\mathbb{Z}_N^{(h)} * x}_H 
        \end{align*}

        \[
        =\sum_{g \in \mathbb{Z}_N} e^{i\lambda_gt} \ket{\mathbb{Z}_N^{(g)} * x} \bra{\mathbb{Z}_N^{(g)} * x} \Big( \frac{1 - i}{2} \ket{\mathbb{Z}_N^{(h)} * x} + \frac{1 + i}{2} \ket{\mathbb{Z}_N^{(-h)} * x} \Big) 
        \]

        \[
        =e^{i\lambda_{h}t} \frac{1 - i}{2} \ket{\mathbb{Z}_N^{(h)} * x} + \frac{1 + i}{2} e^{i\lambda_{-h}t} \ket{\mathbb{Z}_N^{(-h)} * x} 
        \]

        \[
        =e^{i\lambda_{h}t} \ket{\mathbb{Z}_N^{(h)} * x}_H,
        \]

        where the last equality follows from the fact that $\lambda_h = \lambda_{-h}$.
    \end{textblock*}

@@

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        Let $\ket{a} = \ket{t}\ket{b}$, where $b$ is the least significant bit of $a$, so that $a = 2t + b$. Applying $\qft_{N / 2}$ to the first register, we obtain the state
        \[ \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \omega_N^{2ty} \ket{y} \ket{b}. \]
        Next, we apply the phase unitary $P(y, b): \ket{y} \ket{b} \mapsto \omega_N^{by} \ket{y} \ket{b}$, and finally, we apply a Hadamard transform to the last qubit. The result is the state in \eqref{eq:qft_alt}.

@@

        \begin{align*}
            \ket{0}\ket{t}\ket{b}
            & \mapsto \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi t y}{N / 2} \Big) \ket{0} \ket{y} \ket{b} \\%& (\mathds{1} \otimes \qht_{N / 2} \otimes \mathds{1}) \\
            & = \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{4\pi t y}{N} \Big) \ket{0} \ket{y} \ket{b} \\
            & \mapsto \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{4\pi t y}{N} \Big) (\ket{0} + \ket{1}) \ket{y}\ket{b}. %& (H \otimes \mathds{1}) 
        \end{align*}

>>>>

% Packages
\usepackage{graphicx}    % For including images
\usepackage{amsmath}     % For mathematical symbols
\usepackage{braket}      % For Dirac notation
\usepackage[absolute,overlay]{textpos}
\usepackage{amsmath}
\usepackage{tikz}

% \usepackage{amsmath, amssymb, dsfont}
% \usepackage{algorithm}
% \usepackage{algorithmicx}
% \usepackage{enumitem}

@@

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        In this talk, we explore the interplay between quantum computation, group theory, and cryptography through the lens of quantum walks. The presentation is organized as follows:

        \vspace{0.5cm}
        \begin{enumerate}
            \item \textbf{Quantum Computation Preliminaries} \\
            \vspace{0.1cm}

            \item \textbf{Quantum Walks} \\
            \vspace{0.1cm}

            \item \textbf{Quantum Money Schemes} \\
            \vspace{0.1cm}
            \item \textbf{Verification of our New Quantum Money scheme using Quantum Walks} \\

        \end{enumerate}

@@

\begin{frame}{Example of a Continuous Quantum Walk}
    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        \centering
        \begin{tikzpicture}[scale=1, every node/.style={circle, draw=black, fill=white, minimum size=10pt, inner sep=2pt}]
            % Draw nodes
            \node (-3) at (-3,0) {-3};
            \node (-2) at (-2,0) {-2};
            \node (-1) at (-1,0) {-1};
            \node (0)  at (0,0)  {0};
            \node (1)  at (1,0)  {1};
            \node (2)  at (2,0)  {2};
            \node (3)  at (3,0)  {3};

            % Draw edges
            \draw (-3) -- (-2);
            \draw (-2) -- (-1);
            \draw (-1) -- (0);
            \draw (0)  -- (1);
            \draw (1)  -- (2);
            \draw (2)  -- (3);

            % Add fading dashed arrows to suggest infinite continuation
            \draw[dashed, ->] (-3,0) -- (-3.8,0);
            \draw[dashed, ->] (3,0) -- (3.8,0);
        \end{tikzpicture}

        \begin{itemize}
        \vspace{0.5cm}
        \item \textbf{Graph:} Infinite line with vertices labeled by integers \( n \in \mathbb{Z} \).
        \item \textbf{Adjacency Matrix:} Each vertex is connected to its neighbors \( n \pm 1 \).
        \[
        A_{i,j} = \begin{cases}
            1 & \text{if } |i - j| = 1 \\
            0 & \text{otherwise}
        \end{cases}
        \]

        \item \textbf{Hilbert Space:} Spanned by basis states \( |n\rangle \), representing positions.
        \item \textbf{Time Evolution:} Governed by Schrödinger equation
        \[
        i \frac{d}{dt} |\psi(t)\rangle = A |\psi(t)\rangle
        \quad \Rightarrow \quad
        |\psi(t)\rangle = e^{-iAt} |\psi(0)\rangle
        \]
    \end{itemize}

@@

        \begin{itemize}
            \item \textbf{Initial State:} Particle starts at the origin:
            \[
            |\psi(0)\rangle = |0\rangle
            \]

            \item \textbf{Evolved State:}
            \[
            |\psi(t)\rangle = \sum_{n \in \mathbb{Z}} \psi_n(t) |n\rangle
            \quad \text{where } \psi_n(t) = i^n J_n(2t)
            \]

            \item \textbf{Key Features:}
            \begin{itemize}
                \item Probability: \( |\psi_n(t)|^2 \)
                \item Wave-like, oscillatory distribution due to interference
                \item Faster spread than classical walk: standard deviation grows linearly in \( t \)
                \item No coin space needed — evolution depends only on graph structure
            \end{itemize}
        \end{itemize}

@@

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        \begin{itemize}
            \item \textbf{Hilbert Space:} $\mathcal{H} = \mathcal{H}_C \otimes \mathcal{H}_P$
            \begin{itemize}
                \item $\mathcal{H}_C$: 2D coin space with basis $\{|0\rangle, |1\rangle\}$
                \item $\mathcal{H}_P$: Infinite-dimensional position space with basis $\{|n\rangle : n \in \mathbb{Z}\}$
            \end{itemize}

            \item \textbf{Coin Operator:} Apply a unitary $C$ (e.g., Hadamard) to $\mathcal{H}_C$
            \[
            H = \frac{1}{\sqrt{2}} \begin{bmatrix}
                1 & 1 \\
                1 & -1
            \end{bmatrix}
            \]

            \item \textbf{Shift Operator:} Moves the walker based on coin state:
            \[
            S|0\rangle|n\rangle = |0\rangle|n + 1\rangle, \quad
            S|1\rangle|n\rangle = |1\rangle|n - 1\rangle
            \]

            \item \textbf{One Step:} Apply the unitary operator
            \[
            U = S \cdot (C \otimes I)
            \]
        \end{itemize}

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        \begin{itemize}
            \item \textbf{Initial State:} $|\psi(0)\rangle = |0\rangle \otimes |0\rangle$

            \item \textbf{After One Step:}
            \[
            |\psi(1)\rangle = \frac{1}{\sqrt{2}} \left( |0\rangle|1\rangle + |1\rangle|-1\rangle \right)
            \]

            \item \textbf{Key Properties:}
            \begin{itemize}
                \item Superposition leads to \textbf{parallel exploration} of paths.
                \item Repeated application of $U$ creates \textbf{interference patterns}.
                \item Spread is faster than classical: $\sigma(t) \sim t$ vs. $\sqrt{t}$.
            \end{itemize}

            \vspace{0.5cm}
            \item \textbf{Measurement:}
            \begin{itemize}
                \item Measuring after each step yields a classical random walk.
                \item \textbf{Quantum behavior requires delaying measurement.}
            \end{itemize}
        \end{itemize}

@@

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        \vspace{0.5cm}

        \begin{itemize}
            \item \textbf{CTQW}: Easier to analyze due to direct spectral decomposition.
            \vspace{0.3cm}
            \item \textbf{CTQW}: Harder to implement on quantum circuits (requires simulating \( e^{-iAt} \)).
            \vspace{0.3cm}
            \item \textbf{DTQW}: More complex to analyze (involves coin and shift operators).
            \vspace{0.3cm}
            \item \textbf{DTQW}: Easier to implement on gate-based quantum hardware.
        \end{itemize}

@@

        \begin{figure}[h!]
            \centering
            \begin{tikzpicture}[shorten >=1pt, auto, node distance=2cm,
                thick, main node/.style={circle,draw,font=\sffamily\small\bfseries}]

                % Define nodes explicitly
                \node[main node] (0) at (90:2cm) {0};
                \node[main node] (1) at (45:2cm) {1};
                \node[main node] (2) at (0:2cm) {2};
                \node[main node] (3) at (-45:2cm) {3};
                \node[main node] (4) at (-90:2cm) {4};
                \node[main node] (5) at (-135:2cm) {5};
                \node[main node] (6) at (180:2cm) {6};

                % Node N-1 explicitly
                \node[main node] (N-1) at (135:2cm) {N-1};

                % Ellipsis between node 6 and N-1
                \node at (157.5:2cm) {\dots};

                % Draw edges explicitly
                \draw (0) -- (1) -- (2) -- (3) -- (4) -- (5) -- (6);
                \draw (6) -- (157.5:1.8cm); % connecting dots visually
                \draw (N-1) -- (0); % close the cycle

            \end{tikzpicture}
            \caption{Cayley graph of $\mathbb{Z}_N$ with generators $\{\pm 1\}$}
            \label{fig:cyclic-graph}
        \end{figure}

@@

        \begin{figure}[H]
            \centering
            \begin{tikzpicture}[thick, 
                main node/.style={circle, draw, font=\sffamily\small\bfseries},
                adjacent node/.style={circle, draw, font=\sffamily\small},
                scale=1]

                % Central node g
                \node[main node] (g) at (0,0) {$g$};

                % Define number of adjacent vertices
                \def \K {5}
                \def \radius {2cm}
                \def \angle {360/\K}

                % Draw adjacent nodes
                \foreach \i in {1,...,\K}{
                    \node[adjacent node] (q\i) at ({90+\angle*(\i-1)}:\radius) {$g*q_{\i}$};
                    \draw (g) -- (q\i);
                }

                % Optional dots if there are many vertices (e.g., q_k)
                \node at ({90+\angle*(\K)}:\radius) {\dots};

            \end{tikzpicture}
            \caption{Cayley graph of a group $G$ with generators $\{q_1, q_2, ... , q_k\}$}

            \label{fig:group-action-orbit}
        \end{figure}

@@

@@

@@

        \begin{itemize}
        \item $\ver(h, \ket{\psi})$. First, check whether $\ket{\psi}$ has support in $X_\lambda$. If not, return $0$. Then, apply \textit{cmpIndex} to the state $\ket{\psi}\ket{0}$, and measure the second register to obtain some $h' \in G_\lambda$. If $h' = h$, return $1$; otherwise return $0$.

        \vspace{1.5cm}
        cmpIndex Algorithm:
        \vspace{0.3cm}
        \item  Given a state $\ket{G^{(h)} * x}$, there is an efficient algorithm for computing $h$.
        Specifically, there is a unitary operator that performs the transformation $\ket{G^{(h)} * x} \ket{0} \mapsto \ket{G^{(h)} * x} \ket{h}$:
        \end{itemize}

@@

@@

@@

        \begin{align*}
            \ket{0}\ket{a} = \ket{0}\ket{t}\ket{b}
            & \mapsto \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{2\pi t y}{N / 2} \Big) \ket{0} \ket{y} \ket{b} \\%& (\mathds{1} \otimes \qht_{N / 2} \otimes \mathds{1}) \\
            & = \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{4\pi t y}{N} \Big) \ket{0} \ket{y} \ket{b} \\
            & \mapsto \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \cas\Big( \frac{4\pi t y}{N} \Big) (\ket{0} + \ket{1}) \ket{y}\ket{b}. %& (H \otimes \mathds{1}) 
        \end{align*}
